Lesson 3
========================================================

***
### Reading in Data
```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

***



### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month)

```

***

### Histogram of Friend Count
```{r}
ggplot(data = pf, aes(x = friend_count))+ 
  geom_histogram(binwidth = 25) +
    scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```


***

### Histogram of Friend Count by Gender
```{r}
ggplot(data = na.omit(pf), aes(x = friend_count))+ 
  geom_histogram(binwidth = 25) +
    scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~ gender, ncol = 1)
```


### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```


***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure/365), data = pf) +
       xlab('Number of years using Facebook') +
       ylab('Number of users in sample') +
   geom_histogram(binwidth = .25, color = 'black', fill = '#099DD9') +
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
ggplot(aes(x = age), data = pf) +
       xlab('Age in years') +
       ylab('Number of users in sample') +
   geom_histogram(binwidth = 1, color = 'white', fill = '#000000') +
  scale_x_continuous(limits = c(0,120))
summary(pf$age)
```

***

### Transforming Data
```{r Friend Count Histogram}
summary(pf$friend_count)

library(gridExtra)

p1 = ggplot(aes(x = friend_count), data = pf) +
      geom_histogram() +
      labs(x = "Friend Count", y = "Number of users")
p2 = p1 + scale_x_log10()
p3 = p1 +scale_x_sqrt()

grid.arrange(p1,p2,p3, ncol<=1)
```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
ggplot(aes(x = friend_count), data = pf) +
      geom_histogram() +
      scale_x_log10()
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
      geom_histogram() +
      scale_x_log10() +
      facet_wrap(~gender)

ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 100)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
summary(pf$www_likes)
ggplot(aes(x = www_likes, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(1, 150), breaks = seq(1, 150, 5)) + 
  xlab('Likes on the web') + 
  ylab('Percentage of users with that like count')

ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender)) + 
  scale_x_log10()

by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom='boxplot') +
      coord_cartesian(ylim=c(0,250))

by(pf$friend_count, pf$gender, summary)
```
***

#### On average, who initiated more friendships in our sample: men or women?

```{r Friend Requests by Gender}
by(pf$friendships_initiated, pf$gender, summary)
```

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```


***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!