R-EDA Problemset 4
========================================================

***
```{r}
library(ggplot2)
```

### price vs. size
create a scatterplot of price vs x. using the ggplot syntax
```{r}
ggplot(aes(x = carat, y = price),
       data = diamonds) +
  geom_point()
```
Notes: There seems to be an exponential correlation between price and size (carat).

***

### correlations
what is the correlation between price and x,y,z?
```{r}
cor.test(diamonds$price, diamonds$x, method = "pearson")
cor.test(diamonds$price, diamonds$y, method = "pearson")
cor.test(diamonds$price, diamonds$z, method = "pearson")
```
***

### price vs. depth
create a scatterplot of price vs depth using the ggplot syntax
```{r}
ggplot(aes(x = price, y = depth),
       data = diamonds) +
  geom_point()
```
### adjustments price vs. depth
Change the code to make the transparency of the
points to be 1/100 of what they are now and mark
the x-axis every 2 units. See the instructor notes
for two hints.
```{r}
ggplot(aes(x = depth, y = price),
       data = diamonds) +
  geom_jitter(alpha = 1/100) +
  geom_point() +
  scale_x_continuous(breaks = seq(43,79,2))

range(diamonds$depth)
```
### correlation price vs. depth

```{r}
cor.test(diamonds$depth, diamonds$price, method = "pearson")
```

### price vs. carat
Create a scatterplot of price vs carat
and omit the top 1% of price and carat
values.
```{r}
ggplot(aes(x = carat, y = price),
       data = diamonds) +
  geom_point() +
  xlim(c(0, quantile(diamonds$carat,0.99))) +
  ylim(c(0, quantile(diamonds$price,0.99)))
```
### price vs. volume
Create a scatterplot of price vs. volume (x * y * z).
This is a very rough approximation for a diamond's volume.
Create a new variable for volume in the diamonds data frame.
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot(aes(x = volume, y = price),
       data = diamonds) +
  geom_point()

detach("package:plyr", unload=TRUE)

count(diamonds$volume == 0)
```

### correlation between price and volume
```{r}
new_diamonds <- subset(diamonds, diamonds$volume < 800 & diamonds$volume > 0)

cor.test(new_diamonds$price, new_diamonds$volume, method = "pearson")
```

### adjustments - price vs. volume
# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)

# Do you think this would be a useful model to estimate
# the price of diamonds? Why or why not?
The model could be usefull to predict prices of diamonds with a value between 0 and 350. After this, geom smooth shows that the high variance leads to a dropping price of diamonds larger than 350.
```{r}
ggplot(aes(x = volume, y = price),
       data = new_diamonds) +
  geom_point() +
  geom_smooth()

?geom_smooth
```

# Use the function dplyr package
# to create a new data frame containing
# info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following
# variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each
# level of clarity.

```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n ())

head(diamondsByClarity, 20)
```
# Your task is to write additional code to create two bar plots
# on one output image using the grid.arrange() function from the package
# gridExtra.
```{r}
library(dplyr)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
diamonds_by_clarity_hist <- ggplot(diamonds_mp_by_clarity, aes(x= clarity, y= mean_price)) +
  geom_bar(stat = "identity")

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
diamonds_mp_by_color_hist <- ggplot(diamonds_mp_by_color, aes(x= color, y= mean_price)) +
  geom_bar(stat = "identity")
library(gridExtra)
grid.arrange(diamonds_by_clarity_hist, diamonds_mp_by_color_hist)
```

