Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

People guess whole numbers -> horizonzal lines
Most points fall below the perfect accuracy line, with a larger actual audience than percived audience

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('/Users/lt/Git/R-EDA/lesson3/pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:
Young users have higher friend counts than older ones. There are peeks at certain ages such as ~70 and ~100 which should be investigated.
***

### ggplot Syntax
Notes:
- Building up plots one layer at a time!
- The distribution of users with a lot of friends has multiple peaks, the by far largest between 13-25 and another one around 50

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13,90) 
  
summary(pf$age)
```

***

### Overplotting
Notes:
Jitter needs to be removed as negative numbers can be produced by the noise function which would cause errors in the sqrt function.
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```

#### What do you notice in the plot?
Response:
For ages above 65 the number of friends has a smaller range, there are fewer users with 0 friends.

***

### Alpha and Jitter
Explore the relationship between friends initiated vs. age
Notes: There seem to be two peeks for friendships initiated. One is for users between 13-25 and the other for users ~50 years. There seems to be an abnormally amount of users with age 70 that initiated a lot of friendships.

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = 'sqrt')
```


***

### Conditional Means
Notes:
# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.
```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n ()) %>%
  arrange(age)

head(pf.fc_by_age, 20)

ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()+
  geom_point()

```


### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) +
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = 'orange') +
   coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile,  fun.args = list(probs = .1),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile,  fun.args = list(probs = .9),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = median, color = 'blue')
```

#### What are some of your observations of the plot?
Response:
The difference between the number of friends and age has a peek with large distances between the quantiles as well as median&mean until ~age 25. From age 25 until age 67 the number of friends is constant, will a small differences between the median&mean and some noise around quantile .1 and .9. Above age 67 the data is very sparse and gets larger distances between quantiles, mean and median.
***


### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = "pearson")
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.1624741 
***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
?subset
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

## Create Scatterplots
Notes:Create a scatterplot of likes_received (y)
vs. www_likes_received (x). Use any of the
techniques that you've learned so far to
modify the plot.

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  xlim(c(0, quantile(pf$www_likes_received,0.95))) +
  ylim(c(0, quantile(pf$likes_received,0.95))) +
  geom_smooth(method = 'lm', color = 'red')
```

***

### Strong Correlations
What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Strong Correlations}
cor.test(pf$likes_received, pf$www_likes_received, method = "pearson")
```


```{r Correlation Calcuation
### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red')
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
~0
b. What is the actual correlation of the two variables?
(Round to the thousandths place)
0.057

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp, method = "pearson")
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  scale_x_continuous(breaks = seq(0,203,12))
```

***

### A New Perspective

What do you notice?
Response:
The unique month and temperature have a correlation. Summer has higher temperatures than winter.
```{r}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+ 
  geom_point()
```



### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (1 - pf$dob_month / 12) 
```

***

### Age with Months Means

```{r Age with Months Means}
library(dplyr)
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n ()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months, 20)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean),
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age_with_months, y = friend_count_mean),
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age, y = friend_count_mean),
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()

library(gridExtra)
grid.arrange(p2,p1,ncol = 1)
```

***

### Which Plot to Choose?
Notes:
No need to chose - multiple plots can and should be created. They might reveal different insights on the dataset. When presenting it should be boiled down to only those important charts.

***

### Analyzing Two Variables
Reflection:
- scatterplots
- conditional means
- correlation coefficients
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

